Type-Safe Modular Parsing

### 2017-01-08 paper/sections/Inheritance.tex
Organise code into sections, some fixes
Code

\lstinputlisting[linerange=30-32]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_SIMPLE_LAM

@@

\huang{Extending from old parser is necessory to get the `pExpr', added a surrounding trait.}
\lstinputlisting[linerange=43-46]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_BAD_ATTEMPT

@@

\lstinputlisting[linerange=50-55]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_APPROACH
Now \inlinecode{VarExprParser} successfully represents the parser for the extended language, because Scala uses dynamic dispatch for
method overriding in inheritance. When the input \inlinecode{"1 + x"} is fed to the parser \inlinecode{this.pExpr}, it firstly delegates
the work to \inlinecode{super.pExpr}, which parses literals and additions. However, the recursive call \inlinecode{pExpr} in \inlinecode{pAdd}
actually refers to \inlinecode{this.pExpr} again, due to dynamic dispatch. It implies that the recursive call has included the extension. This
elegant inheritance-based approach opens the recursion of parsers in a modular way.

@@

\lstinputlisting[linerange=60-66]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE

>>>>

\lstinputlisting[linerange=29-31]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_SIMPLE_LAM

@@

\huang{Extending from old parser is necessory to get the `pExpr', added a surrounding trait.}
\lstinputlisting[linerange=35-38]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_BAD_ATTEMPT

@@

\lstinputlisting[linerange=42-47]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_APPROACH
Now \inlinecode{VarExprParser} successfully represents the parser for the extended language, because Scala uses dynamic dispatch for
method overriding in inheritance. When the input \inlinecode{"1 + x"} is fed to the parser \inlinecode{this.pExpr}, it firstly delegates
the work to \inlinecode{super.pExpr}, which parses literals and additions. However, the recursive call \inlinecode{pExpr} in \inlinecode{pAdd}
actually refers to \inlinecode{this.pExpr} again, due to dynamic dispatch. It implies that the recursive call has included the extension. This
elegant inheritance-based approach opens the recursion of parsers in a modular way.

@@

\lstinputlisting[linerange=56-62]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE

### 2017-01-08 paper/sections/ObjectAlgebras.tex
Code

\lstinputlisting[linerange=9-17]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
Such a parser works fine, yet it does not appear to be extensible. For example, we have demonstrated in Section~\ref{subsec:inheritance-approach} that method overriding can update \inlinecode{pExpr} for an extended syntax, but now \inlinecode{pExpr} will have a different type, like \inlinecode{VarExprAlg[E] =>} \inlinecode{Parser[E]} where the difference appears in the argument. The difference of types prohibits method overriding, however, this problem refers to \textit{binary methods}. \haoyuan{Shall we mention this?}

One solution to this problem is taking out the argument as a single field, hence it supports overriding as well. Therefore we refactor the parsing code as follows:
\lstinputlisting[linerange=24-34]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER
This is precisely the pattern that we advocate for modular parsing. Note that \inlinecode{pE} is the one that is used recursively, exported by the trait and considered to be overridden from inheritance. The reason why we use an extra and seemingly redundant \inlinecode{pExpr}, is that \inlinecode{pE} can no longer be referred to by the \inlinecode{super} keyword, as a limitation of Scala for generics, hence we can directly invoke \inlinecode{pExpr} as the specific \inlinecode{pE} in \inlinecode{ExprOAParser}. And we still want code reuse for the composition of \inlinecode{pLit} and \inlinecode{pAdd}. Now it is worth trying on this pattern with an extension, like \inlinecode{VarExprAlg} which we have been using:
\lstinputlisting[linerange=38-49]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
In the client code shown above, an instance of algebra is fed to \inlinecode{VarExprOAParser} for pretty-printing expressions. Then the trait exports \inlinecode{pE}, parses the input and applies the algebra. The result shows that such a pattern is a nice solution to modular parsing. At this point, one may argue that with a different operation like evaluation, the redundant code is required to obtain the parser, and it results in parsing the same input twice. This is generally solved in the original Object Algebras paper [] by composing two algebras into one in advance.

@@

\lstinputlisting[linerange=32-60]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
Besides lexing, the trait \inlinecode{TypedLamOAParser} also introduces parsers for types, and the new case for expressions. For the same reason,
we use \inlinecode{pTypedLamT} and \inlinecode{pTypedLamE} as copies of current \inlinecode{pT} and \inlinecode{pE}, recursively. \inlinecode{pT} and \inlinecode{pE} are actually the parsers for recursion.

@@

\lstinputlisting[linerange=64-66]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
Pretty-printing the parsed results just works as expected.

>>>>

\lstinputlisting[linerange=8-16]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER_BAD
Such a parser works fine, yet it does not appear to be extensible. For example, we have demonstrated in Section~\ref{subsec:inheritance-approach} that method overriding can update \inlinecode{pExpr} for an extended syntax, but now \inlinecode{pExpr} will have a different type, like \inlinecode{VarExprAlg[E] =>} \inlinecode{Parser[E]} where the difference appears in the argument. The difference of types prohibits method overriding, however, this problem refers to \textit{binary methods}. \haoyuan{Shall we mention this?}

One solution to this problem is taking out the argument as a single field, hence it supports overriding as well. Therefore we refactor the parsing code as follows:
\lstinputlisting[linerange=23-33]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=BASE_OA_PARSER
This is precisely the pattern that we advocate for modular parsing. Note that \inlinecode{pE} is the one that is used recursively, exported by the trait and considered to be overridden from inheritance. The reason why we use an extra and seemingly redundant \inlinecode{pExpr}, is that \inlinecode{pE} can no longer be referred to by the \inlinecode{super} keyword, as a limitation of Scala for generics, hence we can directly invoke \inlinecode{pExpr} as the specific \inlinecode{pE} in \inlinecode{ExprOAParser}. And we still want code reuse for the composition of \inlinecode{pLit} and \inlinecode{pAdd}. Now it is worth trying on this pattern with an extension, like \inlinecode{VarExprAlg} which we have been using:
\lstinputlisting[linerange=37-48]{../Scala/Parser/src/PaperCode/Sec4OA/Code3.scala}% APPLY:linerange=EXT_OA_PARSER
In the client code shown above, an instance of algebra is fed to \inlinecode{VarExprOAParser} for pretty-printing expressions. Then the trait exports \inlinecode{pE}, parses the input and applies the algebra. The result shows that such a pattern is a nice solution to modular parsing. At this point, one may argue that with a different operation like evaluation, the redundant code is required to obtain the parser, and it results in parsing the same input twice. This is generally solved in the original Object Algebras paper [] by composing two algebras into one in advance.

@@

\lstinputlisting[linerange=31-59]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
Besides lexing, the trait \inlinecode{TypedLamOAParser} also introduces parsers for types, and the new case for expressions. For the same reason,
we use \inlinecode{pTypedLamT} and \inlinecode{pTypedLamE} as copies of current \inlinecode{pT} and \inlinecode{pE}, recursively. \inlinecode{pT} and \inlinecode{pE} are actually the parsers for recursion.

@@

\lstinputlisting[linerange=63-65]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
Pretty-printing the parsed results just works as expected.
